---
title: "Review Questions for HSMP410"
author: "Ted Laderas"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Week 1.2 Introduction to Bioinformatics

1. What is a SNP? What is a variant?
1. What is an Odds Ratio and why do we use it to quantify association?


## Week 2.2 Introduction to Genomics

1. What are some differences between a mutation and a sequencing error?
1. How can a mutation in DNA cause a malfunction in the protein? 

## Week 4.1 Data Literacy and Visualization

Given the following boxplot, what do you conclude? Is there a difference between Sepal Length between species?

```{r}
iris %>% ggplot(aes(x=Species, y=Sepal.Length, fill=Species)) + geom_boxplot()
```

## Week 4.2 Data Science

Given the following dataset, what variables would you need to calculate horsepower divided by miles per gallon?

```{r}
head(mtcars)
```

## Week 6.1 Ethics

A nurse posted on her instagram feed a selfie with her patient, naming names.

Of the four ethical principles (beneficience, non-maleficience, justice, autonomy), take two principles and say whether they are in conflict in this situation.

## Week 7.2 Ontologies and Standards

What are some uses for an ontology?

